---
title: ""
output: html_document
date: "2023-10-28"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# [Assignment 6: Time Series Analysis]{.underline}
Denis Ouellette

#### Loading the necessary packages:

```{r}
library(tidyverse)

library(forecast)

library(tseries)
```

## Loading the Dataset

Steps below convert the data into time series type data for our modelling. The final dataset to be used with the model is 'retailTS1'.

```{r}
setwd("C:/Users/denis/OneDrive/Desktop/Predictive Analysis 420/Datasets")

retailTS <- read.csv("retail_data.csv")

retailTS1 <- select(retailTS,Sales)

retailTS1 <- ts(retailTS1,frequency = 12,start=c(2018,1))

head(retailTS1)
tail(retailTS1)
```

## Creating a decomposition plot:

```{r}
plot(decompose(retailTS1))

```

## Modelling and Forecasting

We will train an ARIMA model with the dataset. First we will smooth the data with a logarithmic transformation.

```{r}
logRetail <- log(retailTS1)

arima_model <- auto.arima(logRetail)
```

Now we can create a 12-month forecast and return the 4,8 and 12 month predictions.

```{r}
retailForecast4 <- forecast(arima_model,4)

retailForecast8 <- forecast(arima_model,8)

retailForecast12 <- forecast(arima_model,12)

print(retailForecast4)

print(retailForecast8)

print(retailForecast12)

plot(retailForecast12)



```
