---
title: ""
output: html_document
date: "2023-10-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Assignment 4 - Linear Regression with R

Denis Ouellette

### Loading Necessary Packages

```{r}
library(tidyverse)

```

### Uploading the Datasets:

```{r}
setwd("C:/Users/denis/OneDrive/Desktop/Predictive Analysis 420/Datasets")
CoffeeBloodPressure <- read.csv("coffee_bloodPressure.csv")
DentalData <- read.csv("dentalService.csv")
```

## Question 1: Coffee & Blood Pressure

### Task 1

```{r}
CorrelationCoffee <- as.character(cor(CoffeeBloodPressure$Cups.of.Coffee,CoffeeBloodPressure$Blood.Pressure..Systolic.Pressure.))

CoffeeBloodPressure %>% ggplot(aes(x=Cups.of.Coffee,y=Blood.Pressure..Systolic.Pressure.))+geom_point()+ geom_smooth(method="lm",se=FALSE,color="navy",formula=y~x,fullrange=TRUE)+ scale_x_continuous(limits = c(0,10))+scale_y_continuous(limits=c(100,180))+xlab("Cups of Coffee")+ylab("Blood Pressure")+geom_text(x=5,y=178,label = paste("Correlation =",CorrelationCoffee))


```

### Task 2

There is a positive linear relationship between the Cups of Coffee drank in a day and the resulting blood pressure recorded for the 48 participants in this sample.

A Multiple R of 0.2884 shows a fairly weak correlation.

A Doctor may recommend reducing the amount of coffee you drink in a day to help relieve high blood pressure.

### Task 3

Creating a training model using 75% of dataset (36 observations):

```{r}
#make reproducible
set.seed(1)
#ID column
CoffeeBloodPressure$id <- 1:nrow(CoffeeBloodPressure)
#use 75% of dataset as training set and 25% as test set
training <- CoffeeBloodPressure %>% sample_frac(0.75)

testing  <- anti_join(CoffeeBloodPressure, training, by = 'id')

linearmodel <- lm(Blood.Pressure..Systolic.Pressure. ~ Cups.of.Coffee, data=training)

summary(linearmodel)

training %>% ggplot(aes(x=Cups.of.Coffee,y=Blood.Pressure..Systolic.Pressure.))+geom_point()+ geom_smooth(method="lm",se=FALSE,color="navy",formula=y~x,fullrange=TRUE)+ scale_x_continuous(limits = c(0,10))+scale_y_continuous(limits=c(100,180))+xlab("Cups of Coffee")+ylab("Blood Pressure")
```

Using the model to predict blood pressure values when drinking 6 and 8 cups of coffee:

```{r}
predict(linearmodel,data.frame(Cups.of.Coffee=c(6,8)))
```

The full regression model would predict the following when drinking 6 and 8 cups of coffee:

```{r}
CoffeeBloodPressureRegression <- lm(Blood.Pressure..Systolic.Pressure.~Cups.of.Coffee,data=CoffeeBloodPressure)

predict(CoffeeBloodPressureRegression,data.frame(Cups.of.Coffee=c(6,8)))
```

## Question 2: Dental Services

First we must convert data into numerical in order to plot and run models:

```{r}
DentalData$Income..K.<-as.numeric(str_trim(substring(DentalData$Income..K.,3)))

DentalData$Price<-as.numeric(str_trim(substring(DentalData$Price,3)))
```

#### Task 1 & 2

Creating a training model using 35/49 \~ 71.4% of the observations:

```{r}
#make reproducible
set.seed(1)
#ID column
DentalData$id <- 1:nrow(DentalData)
#use (5/7)~71.4% of dataset as training set and (2/7)% as test set
training1 <- DentalData %>% sample_frac(5/7)

testing1  <- anti_join(DentalData, training, by = 'id')

linearmodel1 <- lm(Income..K. ~ Price, data=training1)

summary(linearmodel1)

training1 %>% ggplot(aes(x=Price,y=Income..K.))+geom_point()+ geom_smooth(method="lm",se=FALSE,color="navy",formula=y~x,fullrange=TRUE)+ylab("Income")

```

There is a strong positive relationship between the price of the mobile dental services and the income generated.

#### Task 3

Using the model to predict the Income at price points of \$50 and \$200:

```{r}
PricePredictions <- predict(linearmodel1,data.frame(Price=c(50,250)))
```

#### Task 4

Using the predictions above to create the new data points and then add them to our previous linear regression scatter plot:

```{r}
Point1 <- as.numeric(c(64.77291,NA,50,NA))
Point2 <- as.numeric(c(357.56893,NA,250,NA))

DentalDataNew <- rbind(training1,Point1,Point2)

DentalDataNew %>% ggplot(aes(x=Price,y=Income..K.))+geom_point()+ geom_smooth(method="lm",se=FALSE,color="navy",formula=y~x,fullrange=TRUE)+ylab("Income")

```
