---
title: ""
output: html_document
date: "2023-10-26"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# [Assignment 5 - k-Means Algorithm]{.underline}

Denis Ouellette

#### Loading the necessary libraries:

```{r include=FALSE}
library(tidyverse)
library(stats)
library(ggfortify)
library(factoextra)
```

## Task 1: Uploading data into a data frame and cleaning.

```{r}
setwd("C:/Users/denis/OneDrive/Desktop/Predictive Analysis 420/Datasets")

MSleepData <- read.csv("mammal_Sleep.csv")

MSleepData <- MSleepData %>% drop_na(vore)
  #dropping NA values

MSleepData <- MSleepData %>% drop_na(sleep_rem)
  #dropping NA values

MSleepData <- MSleepData %>% drop_na(brainwt)
  #dropping NA values
 
VoreCategory <- MSleepData$vore
  #Taking out feature we will try to cluster by

MSleepData1 <- select(MSleepData,c('sleep_total','sleep_rem'))
  #removing incomplete data features to obtain final dataset for modelling. 

MSleepData1 <- scale(MSleepData1)
```

## Task 2: MSleepData1 will be used for k-means clustering.

"VoreCategory" feature will serve as a cluster comparison having 4 distinct categories.

```{r}
head(MSleepData1)
```

## Task 3: Creating a k-means cluster model with 4 centroids.

```{r}
model_km <- kmeans(MSleepData1,centers = 4,nstart=6)

autoplot(model_km,MSleepData1, frame = TRUE)

```

## Task 4: Including the centroids on the chart

```{r}
ModelClusters <- model_km$cluster
  #Cluster labels vector

ModelCentroids <- model_km$centers

print(ModelCentroids)

fviz_cluster(list(data=MSleepData1, cluster = ModelClusters),show.clust.cent = TRUE,labelsize = 1)


```

### Additional Inquiries: "vore" categories correspond with clusters?

#### 1. "vore" categories correspond with clusters?

Using a table, it is possible to assess whether there is a pattern with "vore" classifications and our models' clusters.

```{r}
table(ModelClusters,VoreCategory)
```

There appears to be no pattern...

#### 2. Was four clusters an ideal number to use?

We can use methods such as the "Elbow" or "Silhouette" to explore:

```{r echo=TRUE}
fviz_nbclust(MSleepData1, kmeans, method="wss") + labs(subtitle = "Elbow Method")

fviz_nbclust(MSleepData1, kmeans, method="silhouette") + labs(subtitle = "Silhouette Method")

```

The results suggest 2 or 3 clusters may have been more optimal for the modelling.
