---
title: "Assignment 3"
output:
  word_document: default
  html_document: default
  pdf_document: default
date: "2023-10-20"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Assignment 3 - Using SQL in R

Denis Ouellette

#### Loading SQL Packages:

```{r}
library(RODBC)

library(DBI)

library(odbc)

library(sqldf)

```

#### Load tidyverse

```{r}
library(tidyverse)

#if applicable install.packages("formattable")
library(formattable)
```

#### Setting up database connection to SQL:

```{r}
db_conn <- odbcConnect("Microsoft SQL", rows_at_time = 1)

```

## Query Questions:

#### 1. Which geographic regions generate the most sales?

```{r}
SalesQuery <- "SELECT ST.CountryRegionCode,ST.Name, SH.TerritoryID,ST.SalesLastYear,ST.SalesYTD, SUM(TotalDue) AS GrossSold
FROM Sales.SalesOrderHeader AS SH
JOIN Sales.SalesTerritory AS ST
ON SH.TerritoryID = ST.TerritoryID
GROUP BY ST.CountryRegionCode,SH.TerritoryID,ST.SalesLastYear,ST.SalesYTD,ST.Name
ORDER BY SalesLastYear DESC"

TerritorySalesMetrics <- sqlQuery(db_conn,SalesQuery,stringsAsFactors = FALSE)

TerritorySalesMetrics$SalesLastYear <- currency(TerritorySalesMetrics$SalesLastYear)
TerritorySalesMetrics$SalesYTD <- currency(TerritorySalesMetrics$SalesYTD)
TerritorySalesMetrics$GrossSold <- currency(TerritorySalesMetrics$GrossSold)

print(TerritorySalesMetrics)
```

Sales results by each region for LastYear, YTD figure given and TotalDue on each sales order.

#### 2. What are the most profitable products?

```{r}
ProfitQuery <- "SELECT PP.Name, SD.ProductID , Sum((SD.OrderQty * (SD.UnitPrice-PP.StandardCost))) AS Profit
FROM Sales.SalesOrderDetail AS SD
JOIN Production.Product AS PP
ON SD.ProductID=PP.ProductID
GROUP BY SD.ProductID, PP.Name
ORDER BY Profit DESC"

ProfitByProduct <- sqlQuery(db_conn,ProfitQuery,stringsAsFactors=FALSE)

ProfitByProduct$Profit <- currency(ProfitByProduct$Profit)

head(ProfitByProduct)


```

Profit was calculated by subtracting the UnitCost of each item from the UnitPrice for each Sales order, and multiplying by the quantity.

## Relevant Plots

#### 1.

```{r}
TerritorySalesMetrics %>%  ggplot(aes(x="",y = SalesYTD, fill = Name)) + geom_bar(width=1,stat="identity",colour="black") + geom_text(aes(label = SalesYTD), position = position_stack(vjust=0.5),size=2.5)+theme_void()+ggtitle("SalesYTD")

TerritorySalesMetrics %>%  ggplot(aes(x="",y = SalesLastYear, fill = Name)) + geom_bar(width=1,stat="identity",colour="black") + geom_text(aes(label = SalesLastYear), position = position_stack(vjust=0.5),size=2.5)+theme_void()+ggtitle("Last Year's Sales")


TerritorySalesMetrics %>%  ggplot(aes(x="",y = GrossSold, fill = Name)) + geom_bar(width=1,stat="identity",colour="black") + geom_text(aes(label = GrossSold), position = position_stack(vjust=0.5),size=2.5)+theme_void()+ggtitle("Sold - Totals on All Sales")

```

\*Northeast, Northwest, Southeast and Southwest refer to regions in the U.S.A .

#### 2.

```{r}
TopTenProfitableProducts <- top_n(ProfitByProduct,10,Profit)
options(scipen = 999)

TopTenProfitableProducts %>% ggplot(aes(y=Profit, x=reorder(Name,-Profit),fill=Name))+geom_bar(stat = "identity",show.legend = FALSE)+theme(axis.text.x = element_text(angle = 45, vjust =1, hjust=1))+xlab("Product Name")+scale_y_continuous(limits = c(0,750000))+geom_text(aes(label=Profit),size=2)
```
